% Quick Start Guide: reports package
% Tyler Rinker
% 2-19-13

```{r setup, include=FALSE}
# html5(out.file="replace")     #When finished
# set global chunk options
opts_chunk$set(cache=TRUE)

library(knitr); library(knitcitations)
write.bibtex(c(Francois2013= citation("bibtex"), "Yihui2013" = citation("knitr"), 
    "Boettiger2013" = citation("knitcitations"), "Rinker2013b" = citation("reports"), 
    "Rinker2013a" = citation("qdap")), file = "reports.bib")
bib <- read.bibtex("reports.bib")
```

# Purpose
reports (`r citep(bib[4])`) is a package that assists in writing flexible reports
and presentations.  The package is designed to be used with RStudio,
MikTex/Tex Live, knitr and Pandoc. 

# reports Will Make You...
- Efficient    
- Organized    
- Consistent     
- Happy    

# Getting Started
**Download:**    
* RStudio - [http://www.rstudio.com/](http://www.rstudio.com/)       
* MikTex - [http://miktex.org/](http://miktex.org/) or    
* Tex Live - [http://www.tug.org/texlive/](http://www.tug.org/texlive/)   
* Pandoc - [http://johnmacfarlane.net/pandoc/](http://johnmacfarlane.net/pandoc/)     
* knitr (`r citep(bib[2])`) - [http://yihui.name/knitr/](http://yihui.name/knitr/)     
* knitcitations (`r citep(bib[3])`) 

```{r include=TRUE, eval=FALSE, }   
install.packages("knitr", "knitcitations")
```


# Installation
[https://github.com/trinker/reports](https://github.com/trinker/reports)    
```{r inst, include=TRUE, eval=FALSE, size='small'}
library(devtools)         
install_github("reports", "trinker") 
```

**note- The user may need to install [qdap](http://cran.r-project.org/web/packages/qdap/index.html) before installing.*

# Set Up .Rprofile
The user can add these options to their .Rprofile:   

`options(bib.loc = "C:/Users/trinker/Desktop/PhD Program/MASTER.bib")`    

`options(name_reports = "Tyler Rinker\\\\University at Buffalo\\\\Department of Learning and Instruction")`     


# Create Report/Paper Project
Now the user is ready to generate a new report/paper project.  Use:    
```{r proj, include=TRUE, eval=FALSE}
setwd(desired.location)   
new_project("NEW")
```

# Get Started!

[Project Directory Workflow](https://dl.dropbox.com/u/61803503/report_directory_guide.pdf)

# References
```{r, echo=FALSE, results='asis'}
bibliography("html") 
```

